Differences from our PPoPP Conference Paper \cite{KBP16}

### 2019-01-15 latex/toms/algorithm.tex
dim tree alg
Included descriptions of ADMM and Nesterov methods

\begin{align}
\M{X} \leftarrow \underset{\M{X} \ge 0}{\argmin} \left\lVert \M{A}\M{X} -\M{B}\right\rVert^2_F
\label{eqn:nls}
\end{align}

@@

\subsubsection{Alternating Direction Method of Multipliers (ADMM)}

\subsubsection{Nesterov-type algorithm}

>>>>

\begin{align}
\M{X} \leftarrow \underset{\M{X} \ge \M{0}}{\argmin} \left\lVert \M{A}\M{X} -\M{B}\right\rVert^2_F
\label{eqn:nls}
\end{align}

@@

\subsubsection{Alternating Direction Method of Multipliers (ADMM)}
In the ADMM solver the optimization problem~\Cref{eqn:nls} is reformulated by introducing an auxiliary variable $\M{\hat{X}}$ of size $k \times n$,
\begin{align}
\begin{split}
&\underset{\M{X},\M{\hat{X}}}{\min} \quad \frac 1 2 \left\lVert \M{A}\M{\hat{X}} - \M{B}\right\rVert_F^2 + r(\M{X}) \\
&\textnormal{subject to } \M{X} = \M{\hat{X}}^\Tra 
\end{split}
\label{eqn:admmnls}
\end{align}
Here $r(\cdot)$ is the indicator function of $\mathbb{R}_+$. The updates for the ADMM algorithm becomes,
\begin{align}
\begin{split}
\M{\hat{X}} &\leftarrow \left(\M{A}^\Tra\M{A} + \rho\M{I} \right)^{-1}\left(\M{A}^\Tra\M{B} + \rho\left(\M{X} + \M{U}\right)^\Tra\right) \\
\M{X} &\leftarrow \underset{\M{X}}{\argmin}\ r(\M{X}) + \frac{\rho}{2} \left\lVert \M{X} - \M{\hat{X}}^\Tra + \M{U}\right\rVert_F^2 \\
\M{U} &\leftarrow \M{U} + \M{X} - \M{\hat{X}}^\Tra
\end{split}
\end{align}
where $\M{U}$ is the scaled version of the dual variables corresponding to the equality constraints $\M{X} = \M{\hat{X}}^\Tra$ and $\rho$ is a step size specified by the user. The advantage of using ADMM is the clever splitting of the non-negativity constraints into updates of two blocks of variables $\M{X}$ and $\M{\hat{X}}$. This allows for an unconstrained least squares solve for $\M{\hat{X}}$ and element-wise projections onto $\mathbb{R}_+$ for $\M{X}$. One important fact to notice is that the same matrix $\M{A}^\Tra\M{B}$ and matrix inverse $\Br{\M{A}^\Tra\M{A} + \rho \M{I}}^{-1}$ is used for all the updates. We can therefore cache $\M{A}^\Tra\M{B}$ and the Cholesky decomposition of  $\Br{\M{A}^\Tra\M{A} + \rho \M{I}}$ to save some computations. A comprehensive guide to the ADMM method, convergence properties and selection of optimal $\rho$ can be found in~\cite{boyd2011distributed}.

\subsubsection{Nesterov-type algorithm}
The Nesterov-type algorithm in this package was introduced by Liavas et al~\cite{LKLHS2017}. Their method solves a modified version of NLS problem~\Cref{eqn:nls} with the introduction of a proximal term. The proximal term is useful to handle ill-conditioned instances and guarantee strong convexity. The objective function tackled is,
\begin{align}
\underset{\M{X} \ge \M{0}}{\min} f_p(\M{X}) := \frac 1 2 \left\lVert \M{B}^\Tra - \M{X}^\Tra\M{A}^\Tra \right\rVert_F^2 + \frac{\lambda}{2}\left\lVert\M{X} -\M{X_*}\right\rVert_F^2
\label{eqn:nesnls}
\end{align}
The gradient of $f_p$ is given by the expression,
\begin{align}
\nabla f_p(\M{X}) = -(\M{B}^\Tra - \M{X}^\Tra\M{A}^\Tra)\M{A} + \lambda(\M{X}-\M{X_*})
\end{align}
The first order updates to $\M{X}$ are iterative and sketched below. They are repeated until a termination criteria is triggered. Different criteria are discussed in~\cite{LKLHS2017},
\begin{align}
\begin{split}
\M{Y}_{k+1} &\leftarrow \left[ \M{X}_k - \alpha \nabla f_p(\M{X}_k)\right]_+ \\
\M{X}_{k+1} &\leftarrow \M{Y}_{k+1} + \beta \Br{\M{Y}_{k+1} - \M{Y}_k}
\end{split}
\label{eqn:nesupd}
\end{align}
Here $[\cdot]_+$ is the projection operator in to $\mathbb{R}_+$. The selection of hyperparameters $\lambda, \alpha$ and $\beta$ depends on the spectrum of $\M{A}\M{A}^\Tra$ and is necessary for developing an Nesterov-like method for solving~\Cref{eqn:nesnls}. Details of the selection procedure and different cases can be found in the original paper~\cite{LKLHS2017}. In addition to the acceleration performed during each NLS solve~\Cref{eqn:nesupd} we can also perform an acceleration step for every outer iteration in~\Cref{alg:Par-NNCP-short}~\Cref{line:while}. In this step all factor matrices are updated using the previous iterate values till the objective stops decreasing.
